Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2025-01-07 article.tex
Updates based on Ivan's review comments

\author*[1,2,3]{\fnm{Philip} \sur{Cardiff}}\email{philip.cardiff@ucd.ie}
\author[4]{\fnm{Ivan} \sur{Batisti\'{c}}}
\author[5]{\fnm{Hiroaki} \sur{Nishikawa}}
\author[4]{\fnm{\v{Z}eljko} \sur{Tukovi\'{c}}}

\affil*[1]{\orgdiv{School of Mechanical and Materials Engineering}, \orgname{University College Dublin}, \orgaddress{\country{Ireland}}}
\affil[2]{\orgdiv{UCD Centre for Mechanics}, \orgname{University College Dublin}, \orgaddress{\country{Ireland}}}
\affil[3]{\orgdiv{SFI I-Form Centre}, \orgname{University College Dublin}, \orgaddress{\country{Ireland}}}
\affil[4]{\orgdiv{Faculty of Mechanical Engineering and Naval Architecture}, \orgname{University of Zagreb}, \orgaddress{\country{Croatia}}}
\affil[5]{\orgname{National Institute of Aerospace}, \orgaddress{\country{Hampton, VA 23666, USA}}}

\abstract
{
This study investigates the efficacy of Jacobian-free Newton-Krylov methods in finite-volume solid mechanics. Traditional Newton-based approaches require explicit Jacobian matrix formation and storage, which can be computationally expensive and memory-intensive.
In contrast, Jacobian-free Newton-Krylov methods approximate the Jacobian’s action using finite differences, combined with Krylov subspace solvers such as GMRES, enabling seamless integration into existing segregated finite-volume frameworks without major code refactoring.
This work benchmarks the performance of a compact-stencil Jacobian-free Newton-Krylov method against a conventional segregated approach on a suite of test cases, encompassing varying geometric dimensions, nonlinearities, dynamic responses, and material behaviours.
Key metrics, including computational cost, memory efficiency, and robustness, are evaluated, along with the influence of preconditioning strategies and stabilisation scaling.
Results show that the proposed Jacobian-free Newton-Krylov method outperforms the segregated approach in all linear and nonlinear elastic cases, achieving order-of-magnitude speedups in many instances; however, divergence is observed in elastoplastic cases, highlighting areas for further development.
It is found that preconditioning choice significantly impacts performance: a LU direct solver is fastest in small to moderately-sized cases, while a multigrid method is more effective for larger problems.
The findings demonstrate that Jacobian-free Newton-Krylov methods are promising for advancing finite-volume solid mechanics simulations, particularly for existing segregated frameworks where minimal modifications enable their adoption.
The described implementations are available in the solids4foam toolbox for OpenFOAM, inviting the community to explore, extend, and compare these procedures.
}

@@

Jacobian-free Newton-Krylov methods circumvent the need for the Jacobian matrix by combining the Newton-Raphson method with Krylov subspace iterative linear solvers, such as GMRES, and noticing that such Krylov solvers do not explicitly require the Jacobian matrix.
Instead, only the action of the Jacobian matrix on a solution-type vector is required.
The key step in Jacobian-free Newton-Krylov methods is the approximation of products between the Jacobian matrix and a vector using the finite difference method; that is
\begin{eqnarray}
	\bb{J} \bb{v} \approx \frac{\bb{F}(\bb{x} + \epsilon \bb{v}) - \bb{F}(\bb{x})}{\epsilon}
\end{eqnarray}
where $\bb{J}$ is the Jacobian matrix, $\mathbf{x}$ is the current solution vector (e.g. nodal displacements), $\mathbf{v}$ is a vector (e.g., from a Krylov subspace), and $\epsilon$ is a small scalar perturbation.
%Determining the appropriate value for $\epsilon$ requires balancing the truncation error of the finite difference approximation and round-off (numerical precision) error.
With an appropriate choice of $\epsilon$ (balancing truncation and round-off errors), the characteristic quadratic convergence of Newton methods can be achieved without the Jacobian, hence the modifier \emph{Jacobian-free}.
This approach promises significant memory savings over Jacobian-based methods, especially for large-scale problems, but also potentially for execution time, with appropriate choice of solution components.

@@

More generally,  linear momentum conservation can be expressed in a nonlinear geometry form, which is suitable for finite strains.
Two equivalent nonlinear geometry forms are common: the \emph{total} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_TL}
    \int_{\Omega_o} \rho_o \frac{\partial^2 \bb{u} }{\partial t^2} d\Omega_o
    =
    \oint_{\Gamma_o} \left( J \bb{F}^{-T} \cdot \bb{n}_o \right) \cdot \bb{\sigma} \ d\Gamma_o
    + \int_{\Omega_o}  \bb{f}_b \, d\Omega_o
\end{eqnarray}
and the \emph{updated} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_UL}
    \int_{\Omega_u} \frac{\partial }{\partial t} \left( \rho_u \frac{\partial \bb{u} }{\partial t} \right) d\Omega_u
    = \oint_{\Gamma_u}(j\bb{f}^{-T}\cdot{\bb{n}_u)\cdot \bb{\sigma}}\ d\Gamma_u
    + \int_{\Omega_u}  \bb{f}_b \, d\Omega_u
\end{eqnarray}
where subscript $o$ indicates quantities in the initial reference configuration, and subscript $u$ indicates quantities in the updated configuration.
The true (Cauchy) stress tensor is indicated by $\bb{\sigma}$.
The deformation gradient is defined as $\bb{F} = \textbf{I} + (\bb{\nabla} \bb{u})^T$ and its determinant as $J = \text{det}(\bb{F})$.
Similarly, the \emph{relative} deformation gradient is given in terms of the displacement \emph{increment} as $\bb{f}=\textbf{I} + \left[\bb{\nabla}(\Delta \bb{u}) \right]^T$ and its determinant as $j = \text{det}(\bb{f})$.
The displacement increment is the change in displacement between the current time step and the previous time step when the time interval is discretised into a finite number of steps.

%The two forms are connected through Nanson’s formula \cite{bathe_finite_1996}, which relate the deformed area vector $\bb{\Gamma}$ with the initial area vector $\bb{\Gamma}_{o}$:
%\begin{equation}
%    \bb{\Gamma} = J\bb{F}^{-T}\cdot\bb{\Gamma}_o
%\end{equation}
%Although the total Lagrangian approach is a viable option for wire drawing, the current work adopts the updated Lagrangian approach as developing Eulerian-type upstream and downstream conditions (Section \ref{sec:euler_BCs}) is conceptually easier in an updated Lagrangian formulation.

@@

In Newton-type methods, a Taylor expansion about a current point $\bb{u}_k$ can be used to solve Equation \ref{eqn:residual} \cite{Knoll2004}:
\begin{eqnarray}
	\bb{R}(\bb{u}_{k+1}) = \bb{R}(\bb{u}_{k}) \;+\;  \bb{R}'(\bb{u}_{k}) (\bb{u}_{k+1} - \bb{u}_{k}) \;+\; \text{H.O.T.} = \bb{0}
\end{eqnarray}
Neglecting the higher-order terms ($\text{H.O.T.}$) yields the strict Newton method in terms of an iteration over a sequence of linear systems: 
\begin{eqnarray} \label{eq:NewtonRaphson}
	\bb{J}(\bb{u}_k) \delta \bb{u} &=& -\bb{R}(\bb{u}_n), \notag \\
	\bb{u}_{k+1} &=& \bb{u}_k + s \, \delta \bb{u}, \notag \\
	\quad
	k &=& 0,1,...
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
\end{eqnarray}
where $\bb{J} \equiv \bb{R}' \equiv \partial \bb{R}/\partial \bb{u}$ is the Jacobian matrix.
Starting the Newton procedure requires the specification of $\bb{u}_0$.
%is iteratively solved by linearisation about the current value of the solution, leading to a linear system and iterative update of the solution vector:
%\begin{eqnarray}
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
%\end{eqnarray}
%where subscript $n$ indicates the outer (Newton) iteration index.
The scalar $s > 0$ can be chosen to improve convergence, for example, using a line search or under-relaxation/damping procedure, and is equal to unity in the classic Newton-Raphson approach.
Iterations are performed over this system until the residual $\bb{R}(\bb{u}_n)$ and solution correction $\delta \bb{u}$ are sufficiently small, with appropriate normalisation.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Cell-Centred Finite Volume Discretisation}
\label{sec:discretisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this work, a nominally second-order cell-centred finite volume discretisation is employed.
% based on previously described work, for example, \citep{Cardiff2017, Batistic2022, Tukovic2013, Jasak2000}.
%Consequently, only a summary of the discretisation is presented below.
The solution domain is discretised in both space and time.
The total simulation period is divided into a finite number of time increments, denoted as $\Delta t$, and the discretised governing momentum equation is solved iteratively in a time-marching fashion.
%The spatial domain is partitioned into a finite number of contiguous convex polyhedral cells.
The spatial domain is partitioned into a finite set $\mathcal{P}$ of contiguous convex polyhedral cells, where each cell is denoted by $P \in \mathcal{P}$.
The number of cells in the mesh is indicated by $|\mathcal{P}|$.
A representative cell $P$ is shown in Figure~\ref{fig:cell}.
The set of all faces of cell $P$ is denoted by $\mathcal{F}_P$. This set is further subdivided into two disjoint subsets:
\begin{itemize}
    \item \textbf{Internal Faces} ($\mathcal{F}^{\text{int}}_P$): Faces that are shared with neighbouring cells.
    \item \textbf{Boundary Faces} ($\mathcal{F}^{\text{bnd}}_P$): Faces that lie on the boundary of the spatial domain.
    The boundary faces of cell $P$ are further classed into three disjoint sets, $\mathcal{F}^{\text{bnd}}_P \coloneqq \mathcal{F}^{\text{disp}}_P \cup \mathcal{F}^{\text{trac}}_P \cup \mathcal{F}^{\text{symm}}_P$, representing boundary faces where displacement ($\mathcal{F}^{\text{disp}}_P$), traction ($\mathcal{F}^{\text{trac}}_P$) and symmetry ($\mathcal{F}^{\text{symm}}_P$) conditions are prescribed.
       % \quad \text{with} \quad \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{trac}}_P = \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{symm}}_P = \mathcal{F}^{\text{trac}}_P \cap \mathcal{F}^{\text{symm}}_P = \emptyset.$
\end{itemize}
Each internal face $f_i \in \mathcal{F}^{\text{int}}_P$ corresponds to a neighbouring cell $N_{f_i} \in \mathcal{N}_P$, where $\mathcal{N}_P$ is the set of all neighbouring cells of $P$. The outward unit normal vector associated with an internal face $f_i$ is denoted by $\mathbf{n}_{f_i}$, while the outward unit normal vector associated with a boundary face $b_i$ is denoted by $\mathbf{n}_{b_i}$.
The vector $\mathbf{d}_{f_i}$ connects the centroid of cell $P$ with the centroid of the neighbouring cell $N_{f_i}$, whereas the vector $\mathbf{d}_{b_i}$ connects the centroid of cell $P$ with the centroid of boundary face ${b_i}$.
For convenience, we will also define the set of all faces of cell $P$ excluding those on a traction boundary as $\mathcal{F}_P^{\text{non-trac}} \coloneqq \mathcal{F}_P^{\text{int}} \cup \mathcal{F}_P^{\text{disp}} \cup \mathcal{F}_P^{\text{symm}}$.
%The proposed solution discretisation follows closely the approach of \citet{cardiff_lagrangian_2017}; consequently, only an overview of the final discretised form of equations and adopted solution algorithm are given below.
\begin{figure}[htbp]
	\centering
%	\subfigure[Magnitude of the manufactured displacement solution]
%	{
%		\label{fig:mms_solution}
   		\includegraphics[height=0.45\textwidth]{figures/cell} 
%   	}
%	\subfigure[Polyhedral mesh with $1\,000$ cells]
%	{
%		\label{fig:mms_mesh}
%   		\includegraphics[height=0.45\textwidth]{figures/mms_mesh}  
%   	}
	\caption{Representative convex polyhedral cell $P$ and neighbouring cell $N_{f_i}$, which share a face $f_i$.}
	\label{fig:cell}
\end{figure}

@@

Using Equation \ref{eq:volume_integral}, the inertia term (e.g. left-hand side term of Equation \ref{eqn:momentum_lingeom}) becomes
\begin{eqnarray} \label{eq:inertia}
	\int_{\Omega_P} \rho \frac{\partial \bb{u} }{\partial t}  d \Omega_P
%	\;&\approx&\;
%	\int_{\mathrm{\Omega}} \rho \left[ \frac{\partial \bb{u} }{\partial t}_P + (\bb{x} - \bb{x}_P) \cdot \nabla \frac{\partial \bb{u} }{\partial t}_P \right] d\mathrm{\Omega} \notag \\
	\;&\approx&\;
	\rho_P \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P  \Omega_P
\end{eqnarray}
Similarly, the body force term (e.g. the second term on the right-hand side of Equation \ref{eqn:momentum_lingeom}) becomes:
\begin{eqnarray}
	\int_{\Omega_P} \, \rho \, \bb{g} \,  d \Omega_P
	\;&\approx&\;
	\rho_P \, \bb{g}\,  \Omega_P
\end{eqnarray}
%where subscript $P$ indicates a quantity at the cell centre.
The discretisation of the acceleration in Equation \ref{eq:inertia} can be achieved using one of many finite difference schemes, e.g. first-order Euler, second-order backwards, second-order Newmark-beta.
In the current work, the second-order backwards (BDF2) scheme is used:
%\begin{eqnarray}
%	 \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P
%	 &\approx& \frac{3\bb{v}_{t+1} - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t} \notag \\
%	&\approx& \frac{3\left( \frac{3\bb{u}_{t+1} - 4\bb{u}_{t} + \bb{u}_{t-1}}{2\Delta t} \right) - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t}
%\end{eqnarray}
\begin{eqnarray} \label{eq:inertia2}
	\left(\frac{\partial^2 \boldsymbol{u}_P}{\partial t^2}\right)_P
	&\approx& \frac{3\boldsymbol{v}_P^{[t+1]} - 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t} \notag \\
	&\approx&
	\frac{3\left( 
		\frac{3\boldsymbol{u}_P^{[t+1]} - 4\boldsymbol{u}_P^{[t]} + \boldsymbol{u}_P^{[t-1]}}{2\Delta t} 
		\right) 
	- 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t}
\end{eqnarray}
where $\Delta t$ is the time increment -- assumed constant here -- and $\bb{v} = \partial \bb{u}/\partial t$ is the velocity vector.
Superscript $[t]$ indicates the time level, with $\bb{u}_P^{[t+1]}$ corresponding to the unknown displacement at the current time step.
Consequently, the displacement and velocity at the two previous time steps must be stored, or alternatively the displacement at the previous four time steps.

@@

Within the square braces on thr right-hand side of Equation \ref{eq:RhieChow}, the first terms represent a compact stencil (two-node) approximation of the face normal gradient, while the second terms represent a larger stencil approximation.
These two terms cancel out in the limit of mesh refinement (or if the solution varies linearly); otherwise, they produce a stabilisation effect that tends to smooth the solution fields.
The magnitude of the stabilisation is shown in Appendix \ref{app:RhieChow} to reduce at a second-order rate (not a third-order rate as stated previously \citep{Demirdzic1995}), and hence does not affect the overall scheme's second-order accuracy.

@@

\subsection{Linear Elasticity}
The definition of engineering stress $\bb{\sigma}_s$ for linear elasticity can be given as
\begin{eqnarray}  \label{eq:linearElastic}
	\bb{\sigma}_s
	&=& 2 \mu \bb{\varepsilon} + \lambda \, \text{tr} \left( \bb{\varepsilon} \right) \textbf{I} \notag \\
	&=& \mu \bb{\nabla} \bb{u} + \mu \left( \bb{\nabla} \bb{u}\right)^T + \lambda \left(\bb{\nabla} \cdot \bb{u} \right) \textbf{I}
\end{eqnarray}
where $\lambda$ is the first Lam\'{e} parameter, and $\mu$ is the second Lam\'{e} parameter, synonymous with the shear modulus.
The Lam\'{e} parameters can be expressed in term of the Young's modulus ($E$) and Poisson's ratio $\nu$ as
\begin{eqnarray}
	\mu = \frac{E}{2(1 + \nu)}, \quad \lambda = \frac{E \nu}{(1+\nu)(1 - 2\nu)}
\end{eqnarray}

>>>>

\author*[1,2,3]{\fnm{Philip} \sur{Cardiff}}\email{philip.cardiff@ucd.ie}
\author[4]{\fnm{Ivan} \sur{Batisti\'{c}}}
%\author[5]{\fnm{Hiroaki} \sur{Nishikawa}}
\author[4]{\fnm{\v{Z}eljko} \sur{Tukovi\'{c}}}

\affil*[1]{\orgdiv{School of Mechanical and Materials Engineering}, \orgname{University College Dublin}, \orgaddress{\country{Ireland}}}
\affil[2]{\orgdiv{UCD Centre for Mechanics}, \orgname{University College Dublin}, \orgaddress{\country{Ireland}}}
\affil[3]{\orgdiv{SFI I-Form Centre}, \orgname{University College Dublin}, \orgaddress{\country{Ireland}}}
\affil[4]{\orgdiv{Faculty of Mechanical Engineering and Naval Architecture}, \orgname{University of Zagreb}, \orgaddress{\country{Croatia}}}
%\affil[5]{\orgname{National Institute of Aerospace}, \orgaddress{\country{Hampton, VA 23666, USA}}}

\abstract
{
This study investigates the efficacy of Jacobian-free Newton-Krylov methods in finite-volume solid mechanics. Traditional Newton-based approaches require explicit Jacobian matrix formation and storage, which can be computationally expensive and memory-intensive.
In contrast, Jacobian-free Newton-Krylov methods approximate the Jacobian’s action using finite differences, combined with Krylov subspace solvers such as the generalised minimal residual method (GMRES), enabling seamless integration into existing segregated finite-volume frameworks without major code refactoring.
This work benchmarks the performance of a compact-stencil Jacobian-free Newton-Krylov method against a conventional segregated approach on a suite of test cases, encompassing varying geometric dimensions, nonlinearities, dynamic responses, and material behaviours.
Key metrics, including computational cost, memory efficiency, and robustness, are evaluated, along with the influence of preconditioning strategies and stabilisation scaling.
Results show that the proposed Jacobian-free Newton-Krylov method outperforms the segregated approach in all linear and nonlinear elastic cases, achieving order-of-magnitude speedups in many instances; however, divergence is observed in elastoplastic cases, highlighting areas for further development.
It is found that preconditioning choice significantly impacts performance: a LU direct solver is fastest in small to moderately-sized cases, while a multigrid method is more effective for larger problems.
The findings demonstrate that Jacobian-free Newton-Krylov methods are promising for advancing finite-volume solid mechanics simulations, particularly for existing segregated frameworks where minimal modifications enable their adoption.
The described implementations are available in the solids4foam toolbox for OpenFOAM, inviting the community to explore, extend, and compare these procedures.
}

@@

Jacobian-free Newton-Krylov methods circumvent the need for the Jacobian matrix by combining the Newton-Raphson method with Krylov subspace iterative linear solvers, such as the generalised minimal residual method (GMRES), and noticing that such Krylov solvers do not explicitly require the Jacobian matrix.
Instead, only the action of the Jacobian matrix on a solution-type vector is required.
The key step in Jacobian-free Newton-Krylov methods is the approximation of products between the Jacobian matrix and a vector using the finite difference method; that is
\begin{eqnarray}
	\bb{J} \bb{v} \approx \frac{\bb{F}(\bb{x} + \epsilon \bb{v}) - \bb{F}(\bb{x})}{\epsilon}
\end{eqnarray}
where $\bb{J}$ is the Jacobian matrix, $\mathbf{x}$ is the current solution vector (e.g. nodal displacements), $\mathbf{v}$ is a vector (e.g., from a Krylov subspace), and $\epsilon$ is a small scalar perturbation.
%Determining the appropriate value for $\epsilon$ requires balancing the truncation error of the finite difference approximation and round-off (numerical precision) error.
With an appropriate choice of $\epsilon$ (balancing truncation and round-off errors), the characteristic quadratic convergence of Newton methods can be achieved without the Jacobian, hence the modifier \emph{Jacobian-free}.
This approach promises significant memory savings over Jacobian-based methods, especially for large-scale problems, but also potentially for execution time, with appropriate choice of solution components.

@@

More generally,  linear momentum conservation can be expressed in a nonlinear geometry form, which is suitable for finite strains.
Two equivalent nonlinear geometry forms are common: the \emph{total} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_TL}
    \int_{\Omega_o} \rho_o \frac{\partial^2 \bb{u} }{\partial t^2} d\Omega_o
    =
    \oint_{\Gamma_o} \left( J \bb{F}^{-\text{T}} \cdot \bb{n}_o \right) \cdot \bb{\sigma} \ d\Gamma_o
    + \int_{\Omega_o}  \bb{f}_b \, d\Omega_o
\end{eqnarray}
and the \emph{updated} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_UL}
    \int_{\Omega_u} \frac{\partial }{\partial t} \left( \rho_u \frac{\partial \bb{u} }{\partial t} \right) d\Omega_u
    = \oint_{\Gamma_u}(j\bb{f}^{-\text{T}}\cdot{\bb{n}_u)\cdot \bb{\sigma}}\ d\Gamma_u
    + \int_{\Omega_u}  \bb{f}_b \, d\Omega_u
\end{eqnarray}
where subscript $o$ indicates quantities in the initial reference configuration, and subscript $u$ indicates quantities in the updated configuration.
The true (Cauchy) stress tensor is indicated by $\bb{\sigma}$.
The deformation gradient is defined as $\bb{F} = \textbf{I} + (\bb{\nabla} \bb{u})^{\text{T}}$ and its determinant as $J = \text{det}(\bb{F})$.
Similarly, the \emph{relative} deformation gradient is given in terms of the displacement \emph{increment} as $\bb{f}=\textbf{I} + \left[\bb{\nabla}(\Delta \bb{u}) \right]^{\text{T}}$ and its determinant as $j = \text{det}(\bb{f})$.
The displacement increment is the change in displacement between the current time step and the previous time step when the time interval is discretised into a finite number of steps.

%The two forms are connected through Nanson’s formula \cite{bathe_finite_1996}, which relate the deformed area vector $\bb{\Gamma}$ with the initial area vector $\bb{\Gamma}_{o}$:
%\begin{equation}
%    \bb{\Gamma} = J\bb{F}^{-\text{T}}\cdot\bb{\Gamma}_o
%\end{equation}
%Although the total Lagrangian approach is a viable option for wire drawing, the current work adopts the updated Lagrangian approach as developing Eulerian-type upstream and downstream conditions (Section \ref{sec:euler_BCs}) is conceptually easier in an updated Lagrangian formulation.

@@

In Newton-type methods, a Taylor expansion about a current point $\bb{u}_k$ can be used to solve Equation \ref{eqn:residual} \cite{Knoll2004}:
\begin{eqnarray}
	\bb{R}(\bb{u}_{k+1}) = \bb{R}(\bb{u}_{k}) \;+\;  \bb{R}'(\bb{u}_{k}) (\bb{u}_{k+1} - \bb{u}_{k}) \;+\; \text{H.O.T.} = \bb{0}
\end{eqnarray}
Neglecting the higher-order terms ($\text{H.O.T.}$) yields the strict Newton method in terms of an iteration over a sequence of linear systems: 
\begin{eqnarray} \label{eq:NewtonRaphson}
	\bb{J}(\bb{u}_k) \delta \bb{u} &=& -\bb{R}(\bb{u}_k), \notag \\
	\bb{u}_{k+1} &=& \bb{u}_k + s \, \delta \bb{u}, \notag \\
	\quad
	k &=& 0,1,...
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
\end{eqnarray}
where $\bb{J} \equiv \bb{R}' \equiv \partial \bb{R}/\partial \bb{u}$ is the Jacobian matrix.
Starting the Newton procedure requires the specification of $\bb{u}_0$.
%is iteratively solved by linearisation about the current value of the solution, leading to a linear system and iterative update of the solution vector:
%\begin{eqnarray}
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
%\end{eqnarray}
%where subscript $n$ indicates the outer (Newton) iteration index.
The scalar $s > 0$ can be chosen to improve convergence, for example, using a line search or under-relaxation/damping procedure, and is equal to unity in the classic Newton-Raphson approach.
Iterations are performed over this system until the residual $\bb{R}(\bb{u}_k)$ and solution correction $\delta \bb{u}$ are sufficiently small, with appropriate normalisation.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Cell-Centred Finite Volume Discretisation}
\label{sec:discretisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this work, a nominally second-order cell-centred finite volume discretisation is employed.
% based on previously described work, for example, \citep{Cardiff2017, Batistic2022, Tukovic2013, Jasak2000}.
%Consequently, only a summary of the discretisation is presented below.
The solution domain is discretised in both space and time.
The total simulation period is divided into a finite number of time increments, denoted as $\Delta t$, and the discretised governing momentum equation is solved iteratively in a time-marching fashion.
%The spatial domain is partitioned into a finite number of contiguous convex polyhedral cells.
The spatial domain is partitioned into a finite set $\mathcal{P}$ of contiguous convex polyhedral cells, where each cell is denoted by $P \in \mathcal{P}$.
The number of cells in the mesh is indicated by $|\mathcal{P}|$.
A representative cell $P$ is shown in Figure~\ref{fig:cell}.
The set of all faces of cell $P$ is denoted by $\mathcal{F}_P$. This set is further subdivided into two disjoint subsets:
\begin{itemize}
    \item \textbf{Internal Faces} ($\mathcal{F}^{\text{int}}_P$): Faces that are shared with neighbouring cells.
    \item \textbf{Boundary Faces} ($\mathcal{F}^{\text{bnd}}_P$): Faces that lie on the boundary of the spatial domain.
    The boundary faces of cell $P$ are further classed into three disjoint sets, $\mathcal{F}^{\text{bnd}}_P \coloneqq \mathcal{F}^{\text{disp}}_P \cup \mathcal{F}^{\text{trac}}_P \cup \mathcal{F}^{\text{symm}}_P$, representing boundary faces where displacement ($\mathcal{F}^{\text{disp}}_P$), traction ($\mathcal{F}^{\text{trac}}_P$) and symmetry ($\mathcal{F}^{\text{symm}}_P$) conditions are prescribed.
       % \quad \text{with} \quad \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{trac}}_P = \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{symm}}_P = \mathcal{F}^{\text{trac}}_P \cap \mathcal{F}^{\text{symm}}_P = \emptyset.$
\end{itemize}
Each internal face $f_i \in \mathcal{F}^{\text{int}}_P$ corresponds to a neighbouring cell $N_{f_i} \in \mathcal{N}_P$, where $\mathcal{N}_P$ is the set of all neighbouring cells of $P$. The outward unit normal vector associated with an internal face $f_i$ is denoted by $\mathbf{n}_{f_i}$, while the outward unit normal vector associated with a boundary face $b_i$ is denoted by $\mathbf{n}_{b_i}$.
The vector $\mathbf{d}_{f_i}$ connects the centroid of cell $P$ with the centroid of the neighbouring cell $N_{f_i}$, whereas the vector $\mathbf{d}_{b_i}$ connects the centroid of cell $P$ with the centroid of boundary face ${b_i}$.
For convenience, we will also define the set of all faces of cell $P$ excluding those on a traction boundary as $\mathcal{F}_P^{\text{non-trac}} \coloneqq \mathcal{F}_P^{\text{int}} \cup \mathcal{F}_P^{\text{disp}} \cup \mathcal{F}_P^{\text{symm}}$.
%The proposed solution discretisation follows closely the approach of \citet{cardiff_lagrangian_2017}; consequently, only an overview of the final discretised form of equations and adopted solution algorithm are given below.
\begin{figure}[htbp]
	\centering
%	\subfigure[Magnitude of the manufactured displacement solution]
%	{
%		\label{fig:mms_solution}
   		\includegraphics[width=\textwidth]{figures/cell} 
%   	}
%	\subfigure[Polyhedral mesh with $1\,000$ cells]
%	{
%		\label{fig:mms_mesh}
%   		\includegraphics[height=0.45\textwidth]{figures/mms_mesh}  
%   	}
	\caption{Representative convex polyhedral cell $P$ and neighbouring cell $N_{f_i}$, which share a face $f_i$.}
	\label{fig:cell}
\end{figure}

@@

Using Equation \ref{eq:volume_integral}, the inertia term (e.g. left-hand side term of Equation \ref{eqn:momentum_lingeom}) becomes
\begin{eqnarray} \label{eq:inertia}
	\int_{\Omega_P} \rho \frac{\partial \bb{u} }{\partial t}  d \Omega_P
%	\;&\approx&\;
%	\int_{\mathrm{\Omega}} \rho \left[ \frac{\partial \bb{u} }{\partial t}_P + (\bb{x} - \bb{x}_P) \cdot \nabla \frac{\partial \bb{u} }{\partial t}_P \right] d\mathrm{\Omega} \notag \\
	\;&\approx&\;
	\rho_P \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P  \Omega_P
\end{eqnarray}
Similarly, the body force term (e.g. the second term on the right-hand side of Equation \ref{eqn:momentum_lingeom}) becomes:
\begin{eqnarray}
	\int_{\Omega_P} \, \rho \, \bb{g} \,  d \Omega_P
	\;&\approx&\;
	\rho_P \, \bb{g}\,  \Omega_P
\end{eqnarray}
%where subscript $P$ indicates a quantity at the cell centre.
The discretisation of the acceleration in Equation \ref{eq:inertia} can be achieved using one of many finite difference schemes, e.g. first-order Euler, second-order backwards, or second-order Newmark-beta.
In the current work, the second-order backwards (BDF2) scheme is used:
%\begin{eqnarray}
%	 \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P
%	 &\approx& \frac{3\bb{v}_{t+1} - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t} \notag \\
%	&\approx& \frac{3\left( \frac{3\bb{u}_{t+1} - 4\bb{u}_{t} + \bb{u}_{t-1}}{2\Delta t} \right) - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t}
%\end{eqnarray}
\begin{eqnarray} \label{eq:inertia2}
	\left(\frac{\partial^2 \boldsymbol{u}_P}{\partial t^2}\right)_P
	&\approx& \frac{3\boldsymbol{v}_P^{[t+1]} - 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t} \notag \\
	&\approx&
	\frac{3\left( 
		\dfrac{3\boldsymbol{u}_P^{[t+1]} - 4\boldsymbol{u}_P^{[t]} + \boldsymbol{u}_P^{[t-1]}}{2\Delta t} 
		\right) 
	- 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t}
\end{eqnarray}
where $\Delta t$ is the time increment -- assumed constant here -- and $\bb{v} = \partial \bb{u}/\partial t$ is the velocity vector.
Superscript $[t]$ indicates the time level, with $\bb{u}_P^{[t+1]}$ corresponding to the unknown displacement at the current time step.
Consequently, the displacement and velocity at the two previous time steps must be stored, or alternatively the displacement at the previous four time steps.

@@

Within the square braces on the right-hand side of Equation \ref{eq:RhieChow}, the first terms represent a compact stencil (two-node) approximation of the face normal gradient, while the second terms represent a larger stencil approximation.
These two terms cancel out in the limit of mesh refinement (or if the solution varies linearly); otherwise, they produce a stabilisation effect that tends to smooth the solution fields.
The magnitude of the stabilisation is shown in Appendix \ref{app:RhieChow} to reduce at a second-order rate (not a third-order rate as stated previously \citep{Demirdzic1995}), and hence does not affect the overall scheme's second-order accuracy.

@@

\subsection{Linear Elasticity}
The definition of engineering stress $\bb{\sigma}_s$ for linear elasticity can be given as
\begin{eqnarray}  \label{eq:linearElastic}
	\bb{\sigma}_s
	&=& 2 \mu \bb{\varepsilon} + \lambda \, \text{tr} \left( \bb{\varepsilon} \right) \textbf{I} \notag \\
	&=& \mu \bb{\nabla} \bb{u} + \mu \left( \bb{\nabla} \bb{u}\right)^{\text{T}} + \lambda \left(\bb{\nabla} \cdot \bb{u} \right) \textbf{I}
\end{eqnarray}
where $\lambda$ is the first Lam\'{e} parameter, and $\mu$ is the second Lam\'{e} parameter, synonymous with the shear modulus.
The Lam\'{e} parameters can be expressed in term of the Young's modulus ($E$) and Poisson's ratio $\nu$ as
\begin{eqnarray}
	\mu = \frac{E}{2(1 + \nu)}, \quad \lambda = \frac{E \nu}{(1+\nu)(1 - 2\nu)}
\end{eqnarray}
